```{r setup, include=FALSE}
# loading libraries
# general use libraries
# library(tidyverse)
# library(RMySQL)
library(DBI)
library(readr)
library(ggplot2)
library(lubridate)
library(gt)
library(ggvenn)
library(ggridges)
# library(hrbrthemes)
library(gridExtra)
library(ggpubr)
library(tidyr)
library(purrr)
library(tibble)
library(numbers)
library(car)
library(GGally)
library(dotenv)
# libraries for survival curves
library(survival)
library(survminer)
library(broom)
library(xml2)
library(survey) # for cluster-robust chi-squared test with Rao-Scott correction
library(rcompanion) # for calculating Cramer's V of chi-squared tests
library(duckdb) # for handling big data
library(showtext) # for importing and using custom fonts
library(sysfonts)
library(systemfonts)
library(svglite) # for saving plots as svg to preserve custom font
library(lme4) # Poisson GLMMs
library(DHARMa) # GLMM residual diagnostics
library(ineq) # for calculating gini coefficient
library(tidyverse)
library(dplyr) # putting dplyr at the end because for some reason it misbehaves otherwise
library(glmnet) # for LASSO regression 
library(metafor)

# if dplyr is fighting with duckdb when trying to run Step 1 below, restart R session & fully quit R -> run this in console: install.packages(c("dbplyr", "duckdb", "dplyr")) -> then try rerunning this chunk
# or straight in terminal: pkill -f rsession -> Rscript -e 'install.packages(c("dbplyr", "duckdb", "dplyr"), repos="https://cloud.r-project.org")'
knitr::opts_chunk$set(echo = TRUE)
find_repo_root <- function(start = getwd()) {
  cur <- normalizePath(start, mustWork = TRUE)

  repeat {
    if (dir.exists(file.path(cur, "csv_full")) &&
        file.exists(file.path(cur, "filtered_full_sample_repos.csv"))) {
      return(cur)
    }

    parent <- dirname(cur)
    if (identical(parent, cur)) {
      stop("Could not find repo root containing csv_full/")
    }

    cur <- parent
  }
}

repoRoot <- find_repo_root()
csv_full_dir <- file.path(repoRoot, "csv_full", "")
```

# Loading data
Loading data workflow:
In new session: load libraries -> analysis constants setup chunk -> build/load repo-level rq3 table from `csv_full_dir` -> fit repo-level models.

```{r}
## Analysis constants (RUN EVERY SESSION)

#filters established and verified during prelim exploration (see Full_Dataset_Exploration.Rmd doc)
outlier_commit_shas <- c("2b717338b0e02663dfce5142bd636db0b1d3958c","a3f5c98b59b7892866a49ac22ed91cd8c37865d8","038ca6696d9b821b216b8ab06c8dd48aac7750ae","cea8cd84957f4801f3e9ea1f0997fedfbe3c5e35","dc01a470926d8ee353d7627171e0d491bb8e6386","00314a6c4a2a1c422a4cf9a57b15efca1ad0e146","10763bbae6614c3dc569da892ce6453afd0143f4","ec3a5ff1142670bed5f43969581ecd179d2b8d50","66b199403e65ee29364a25cd7b7596e679030e7f","57960940f9a0b6cd9caf0f92c3592be0cad588d5","23a42d5d956b04c49276240875d8dd329eb44642","0fb2e4b1968f9db4f1cf36fe89505f536e82bc94","3760f7ddeba3cd857a4068a5c08df2800c6950ec")

# these are repos we're excluding: statsig-io/docs - because it was deleted so we couldn't collect full data, the rest because according to the data they don't have any ai_coauthored commits
# github/copilot-cli in particular was a separate discover later on, its a cli tool with no source code so we exclueded it for that reason 
excluded_repos <- c("statsig-io/docs", "Daylily-Informatics/daylily-ephemeral-cluster", "TypedDevs/bashunit", "tscircuit/contribution-tracker", "tscircuit/tscircuit.com", "github/copilot-cli")

observation_start_date <- as.POSIXct("2025-05-01 00:00:00", format = "%Y-%m-%d %H:%M:%S", tz = "UTC")
observation_cutoff_date <- as.POSIXct("2026-05-31 23:59:59", format = "%Y-%m-%d %H:%M:%S", tz = "UTC") 
observation_start_year_month <- "2025-05"
observation_cutoff_year_month <- "2026-05" 
calendar_months_in_window <- 13L
rq3_min_language_repos <- 5L
```

```{r}
## Connect to the data (RUN EVERY SESSION)
con <- dbConnect(duckdb())
ll <- tbl(con, paste0("read_parquet('", csv_full_dir, "line_lifecycle.parquet')"))

# commit_classifications at 200 repos is small enough for RAM
commit_classifications <- readr::read_csv(paste0(csv_full_dir, "commit_classifications.csv")) %>%
  mutate(date = lubridate::as_datetime(date),
         primary_operational_intent = as.factor(primary_operational_intent),
         maintenance_class = as.factor(maintenance_class))
# Registering it with DuckDB so joins against ll happen engine-side
duckdb_register(con, "cc", commit_classifications)

# creating yearMonth var for line_lifecycle data
ll <- ll %>% mutate(origin_commit_year_month   = substr(as.character(origin_commit_date), 1, 7), 
                    terminal_commit_year_month = substr(as.character(terminal_commit_date), 1, 7))

# loading other csvs normally
codebase_agent_share_snapshots <- read_csv(paste0(csv_full_dir, "codebase_agent_share_snapshots.csv"))
# casting snapshot_date and snapshot_commit_date as datetime
codebase_agent_share_snapshots$snapshot_date <- as_datetime(codebase_agent_share_snapshots$snapshot_date)
codebase_agent_share_snapshots$snapshot_commit_date <- as_datetime(codebase_agent_share_snapshots$snapshot_commit_date)
codebase_agent_share_source_snapshots <- read_csv(paste0(csv_full_dir, "codebase_agent_share_source_snapshots.csv"))

commit_classifications <- read_csv(paste0(csv_full_dir, "commit_classifications.csv"))
commit_lifecycle_summary <- read_csv(paste0(csv_full_dir, "commit_lifecycle_summary.csv"))

# loading aggregate no review rate PR data
pr_no_review_monthly_metrics <- read_csv(paste0(csv_full_dir, "combined_pr_no_review_monthly_metrics.csv"))
pr_no_review_monthly_metrics <- pr_no_review_monthly_metrics %>% 
  pivot_longer(cols = -repo, names_to = c(".value", "month"), names_pattern = "(.*)_(\\d{4}_\\d{2})") %>% 
  mutate(month = sub("_", "-", month)) %>% rename(year_month = month)
# creating per-repo aggregation 
repo_pr_no_review_aggregate_metrics <-pr_no_review_monthly_metrics %>% group_by(repo) %>% summarise(no_review_rate = sum(no_review_merged_pr_count) / sum(merged_pr_count))
```


## General helpers

```{r}
blank_to_na <- function(x) {
  x <- as.character(x)
  x[!nzchar(x)] <- NA_character_
  x
}

safe_number <- function(value) {
  value <- suppressWarnings(as.numeric(value))
  value[is.na(value)] <- 0
  value
}

format_nullable_number <- function(value) {
  if (length(value) == 0 || is.na(value) || is.nan(value) || is.infinite(value)) {
    return(NA_real_)
  }
  as.numeric(value)
}

weighted_mean_or_na <- function(values, weights) {
  idx <- !is.na(values) & !is.na(weights) & weights > 0
  if (!any(idx)) {
    return(NA_real_)
  }
  stats::weighted.mean(values[idx], weights[idx])
}

test_label <- function(p_value, estimate, higher_label, lower_label) {
  if (is.na(p_value) || is.na(estimate) || p_value >= rq3_alpha) {
    return("no_clear_difference")
  }
  if (estimate > 0) {
    return(higher_label)
  }
  if (estimate < 0) {
    return(lower_label)
  }
  "no_clear_difference"
}

cox_label <- function(p_value, hazard_ratio) {
  if (is.na(p_value) || is.na(hazard_ratio) || p_value >= rq3_alpha) {
    return("no_clear_difference")
  }
  if (hazard_ratio > 1) {
    return("ai_higher_termination_hazard")
  }
  if (hazard_ratio < 1) {
    return("ai_lower_termination_hazard")
  }
  "no_clear_difference"
}

ratio_or_na <- function(numerator, denominator) {
  ifelse(!is.na(denominator) & denominator > 0, numerator / denominator, NA_real_)
}

log1p_or_na <- function(value) {
  ifelse(!is.na(value) & value >= 0, log1p(value), NA_real_)
}

parse_labels <- function(value) {
  if (length(value) == 0 || is.na(value) || !nzchar(trimws(value))) {
    return(character())
  }
  unique(trimws(unlist(strsplit(value, "[;,]"))))
}

collapse_languages <- function(language_values, min_repos = rq3_min_language_repos, requested_reference = "") {
  language_values <- blank_to_na(language_values)
  language_values[is.na(language_values)] <- "unknown"
  counts <- sort(table(language_values), decreasing = TRUE)
  kept_languages <- names(counts[counts >= min_repos])
  grouped <- ifelse(language_values %in% kept_languages, language_values, "other")
  grouped_counts <- sort(table(grouped), decreasing = TRUE)
  levels_ordered <- names(grouped_counts)

  reference <- ""
  if (nzchar(requested_reference) && requested_reference %in% levels_ordered) {
    reference <- requested_reference
  } else if (length(levels_ordered) > 0) {
    reference <- levels_ordered[[1]]
  }
  levels_ordered <- c(reference, sort(setdiff(levels_ordered, reference)))
  factor(grouped, levels = levels_ordered)
}


# cache for expensive fits — delete the .rds in modelsDir to force that model to re-fit
fit_or_load <- function(name, fit_fun) {
  path <- file.path(modelsDir, paste0(name, ".rds"))
  if (file.exists(path)) return(readRDS(path))
  obj <- fit_fun(); saveRDS(obj, path); obj
}

# for loading models already saved we don't want to recompute
load_fit <- function(name) {
  path <- file.path(modelsDir, paste0(name, ".rds"))
  if (!file.exists(path)) stop("no saved fit at ", path)
  readRDS(path)
} 

ai_labels <- c(
  "aider",
  "amazon_q",
  "claude_code",
  "cline",
  "codex",
  "copilot",
  "cursor",
  "devin",
  "gemini",
  "jules",
  "openhands",
  "replit_agent",
  "roo_code",
  "sourcegraph_cody",
  "sweep",
  "windsurf"
)
```

## Build weighted line table from `csv_full_dir`

```{r}
load_line_weighted_for_rq3 <- function() {
  # Prefer the compact weighted table if it already exists in csv_full_dir.
  # This table is much smaller than line_lifecycle because identical line-level
  # survival records are collapsed and represented by a weight.
  line_weighted_path <- file.path(csv_full_dir, "line_weighted.rds")
  if (file.exists(line_weighted_path)) {
    return(readRDS(line_weighted_path))
  }

  # If the compact table is unavailable, build it in memory from line_lifecycle.
  # This fallback does not write any new intermediate files.
  parquet_path <- file.path(csv_full_dir, "line_lifecycle.parquet")
  csv_path <- file.path(csv_full_dir, "line_lifecycle.csv")
  if (file.exists(parquet_path)) {
    line_lifecycle_source <- paste0("read_parquet('", normalizePath(parquet_path, winslash = "/", mustWork = TRUE), "')")
  } else if (file.exists(csv_path)) {
    line_lifecycle_source <- paste0("read_csv_auto('", normalizePath(csv_path, winslash = "/", mustWork = TRUE), "', sample_size = -1)")
  } else {
    stop("missing line_lifecycle.parquet and line_lifecycle.csv in ", csv_full_dir)
  }

  con <- dbConnect(duckdb())
  ll <- tbl(con, line_lifecycle_source)
  cutoff_sql <- sprintf("TIMESTAMP '%s'", format(observation_cutoff_date, "%Y-%m-%d %H:%M:%S"))

  line_weighted <- ll %>%
    # Restrict to the same observation window and exclusions used in the main analysis.
    mutate(origin_commit_year_month = substr(as.character(origin_commit_date), 1, 7)) %>%
    filter(origin_commit_category != "other_bot",
           origin_commit_year_month >= observation_start_year_month,
           origin_commit_year_month <= observation_cutoff_year_month,
           !(origin_commit_sha %in% outlier_commit_shas),
           !(repo %in% excluded_repos)) %>%
    # Survival time is measured from line creation to termination; surviving lines
    # are censored at the observation cutoff.
    mutate(survival_days = sql(sprintf(
      "round(date_diff('microsecond', origin_commit_date,
              CASE WHEN status = 'survived' THEN %s ELSE terminal_commit_date END) / 86400000000.0, 6)", cutoff_sql))) %>%
    filter(survival_days >= 0) %>%
    # Cox models require strictly positive durations, so exact-zero durations get
    # a tiny positive value without changing the substantive time scale.
    mutate(survival_days = if_else(survival_days < 0.000001, 0.0000005, survival_days)) %>%
    # Collapse identical survival records; weight is the number of lines in that cell.
    count(repo, origin_commit_category, origin_commit_year_month, survival_days, status, name = "weight") %>%
    collect()

  dbDisconnect(con, shutdown = TRUE)
  line_weighted
}

line_weighted_rq3 <- load_line_weighted_for_rq3()
```

## Load `csv_full_dir` repo and commit covariates

```{r}
repo_summary <- read_csv(file.path(csv_full_dir, "repo_summary.csv"), show_col_types = FALSE)
# excluding the one repo that's a cli tool with no source code 
repo_summary <- repo_summary %>% filter(repo != "github/copilot-cli")

commit_file_classifications <- read_csv(file.path(csv_full_dir, "commit_file_classifications.csv"))

commit_files <- read_csv(file.path(csv_full_dir, "commit_files.csv"))

sonarqube_snapshot_scans <- read_csv(file.path(csv_full_dir, "sonarqube_snapshot_scans.csv"))
# casting snapshot_date and snapshot_commit_date as datetime
sonarqube_snapshot_scans$snapshot_date <- as_datetime(sonarqube_snapshot_scans$snapshot_date)
sonarqube_snapshot_scans$snapshot_commit_date <- as_datetime(sonarqube_snapshot_scans$snapshot_commit_date)

commits <- read_csv(file.path(csv_full_dir, "commits.csv"), show_col_types = FALSE) %>%
  mutate(date = as_datetime(date))
full_sample_repos <- read_csv(file.path(csv_full_dir, "full_sample_repos.csv"), show_col_types = FALSE) %>%
  transmute(
    repo = repo_slug,
    metadata_repo_url = blank_to_na(repo_url),
    metadata_primary_language = blank_to_na(primary_language),
    metadata_num_stars = as.numeric(num_stars),
    metadata_num_forks = as.numeric(num_forks)
  ) %>%
  distinct(repo, .keep_all = TRUE)

repo_key_lookup <- repo_summary %>%
  distinct(repo_key, repo)

# Convert the weighted line table into the survival format used by per-repo Cox models.
# Each row is a weighted line-survival cell, not an individual physical line.
line_survival <- line_weighted_rq3 %>%
  left_join(repo_key_lookup, by = "repo") %>%
  filter(!is.na(repo_key),
         origin_commit_category %in% c("not_ai_coauthored", "ai_coauthored")) %>%
  mutate(origin_commit_category = factor(origin_commit_category, levels = c("not_ai_coauthored", "ai_coauthored")),
         survival_days_positive = if_else(survival_days > 0, survival_days, 1e-9),
         event = if_else(status == "survived", 0L, 1L),
         weight = as.numeric(weight)) %>%
  filter(!is.na(survival_days_positive),
         survival_days_positive > 0,
         event %in% c(0L, 1L),
         weight > 0)

line_survival %>%
  summarise(
    repos = n_distinct(repo_key),
    rows = n(),
    weighted_lines = sum(weight),
    ai_lines = sum(weight[origin_commit_category == "ai_coauthored"]),
    human_lines = sum(weight[origin_commit_category == "not_ai_coauthored"])
  )
```

# Building Meta Regression Models

## Stage 1 Per-repo Cox models

here we define some constants for filtering:
```{r}
# The Cox fit threshold prevents unusable models for repos with almost no AI
# or human line exposure.
rq3_cox_min_ai_lines <- 100
rq3_cox_min_human_lines <- 100
rq3_alpha <- 0.05
```

```{r}
empty_cox_result <- function(repo, repo_key, reason) {
  data.frame(
    repo = repo,
    repo_key = repo_key,
    status = "skipped",
    reason = reason,
    n_rows = NA_real_,
    weighted_lines = NA_real_,
    ai_lines = NA_real_,
    human_lines = NA_real_,
    ai_events = NA_real_,
    human_events = NA_real_,
    log_hazard = NA_real_,
    hazard_ratio = NA_real_,
    ci_lower = NA_real_,
    ci_upper = NA_real_,
    se = NA_real_,
    p_value = NA_real_,
    concordance = NA_real_,
    classification = "skipped",
    stringsAsFactors = FALSE
  )
}

run_cox_for_repo <- function(repo_rows, repo_key) {
  if (is.null(repo_rows) || nrow(repo_rows) == 0) {
    return(empty_cox_result(repo_key, repo_key, "no_survival_rows"))
  }

  repo_name <- unique(repo_rows$repo)
  repo_name <- if (length(repo_name) > 0) repo_name[[1]] else repo_key

  ai_rows <- repo_rows[repo_rows$origin_commit_category == "ai_coauthored",]
  human_rows <- repo_rows[repo_rows$origin_commit_category == "not_ai_coauthored",]
  ai_lines <- sum(ai_rows$weight)
  human_lines <- sum(human_rows$weight)

  # Skip repos without enough line exposure in both groups to estimate an AI-vs-human contrast.
  if (ai_lines < rq3_cox_min_ai_lines || human_lines < rq3_cox_min_human_lines) {
    result <- empty_cox_result(repo_name, repo_key, "below_minimum_group_lines")
    result$n_rows <- nrow(repo_rows)
    result$weighted_lines <- sum(repo_rows$weight)
    result$ai_lines <- ai_lines
    result$human_lines <- human_lines
    return(result)
  }

  # Cox models need both events and censored observations represented after filtering.
  if (length(unique(repo_rows$event)) < 2) {
    result <- empty_cox_result(repo_name, repo_key, "insufficient_event_variation")
    result$n_rows <- nrow(repo_rows)
    result$weighted_lines <- sum(repo_rows$weight)
    result$ai_lines <- ai_lines
    result$human_lines <- human_lines
    result$ai_events <- sum(ai_rows$event * ai_rows$weight)
    result$human_events <- sum(human_rows$event * human_rows$weight)
    return(result)
  }

  repo_rows$origin_commit_category <- droplevels(repo_rows$origin_commit_category)
  fit_warnings <- character()

  # Within each repo, estimate whether AI-origin lines have a higher or lower
  # termination hazard than non-AI-origin lines. The weights expand each collapsed
  # survival cell back to its line count.
  fit <- tryCatch(
    withCallingHandlers(
      coxph(
        Surv(survival_days_positive, event) ~ origin_commit_category,
        data = repo_rows,
        weights = weight,
        control = coxph.control(timefix = FALSE)
      ),
      warning = function(warn) {
        fit_warnings <<- c(fit_warnings, conditionMessage(warn))
        invokeRestart("muffleWarning")
      }
    ),
    error = function(err) err
  )
  if (inherits(fit, "error")) {
    result <- empty_cox_result(repo_name, repo_key, paste0("coxph_error: ", conditionMessage(fit)))
    result$n_rows <- nrow(repo_rows)
    result$weighted_lines <- sum(repo_rows$weight)
    result$ai_lines <- ai_lines
    result$human_lines <- human_lines
    result$ai_events <- sum(ai_rows$event * ai_rows$weight)
    result$human_events <- sum(human_rows$event * human_rows$weight)
    return(result)
  }

  fit_summary <- summary(fit)
  coef_name <- "origin_commit_categoryai_coauthored"
  if (!(coef_name %in% rownames(fit_summary$coefficients))) {
    result <- empty_cox_result(repo_name, repo_key, "missing_ai_coefficient")
    result$n_rows <- nrow(repo_rows)
    result$weighted_lines <- sum(repo_rows$weight)
    result$ai_lines <- ai_lines
    result$human_lines <- human_lines
    result$ai_events <- sum(ai_rows$event * ai_rows$weight)
    result$human_events <- sum(human_rows$event * human_rows$weight)
    return(result)
  }

  coef_row <- fit_summary$coefficients[coef_name,]
  ci_row <- fit_summary$conf.int[coef_name,]
  # pulling robust se if possible 
  se_value <- if ("robust se" %in% colnames(fit_summary$coefficients)) {
    coef_row[["robust se"]]
  } else {
    coef_row[["se(coef)"]]
  }
  hazard_ratio <- format_nullable_number(ci_row[["exp(coef)"]])
  p_value <- format_nullable_number(coef_row[["Pr(>|z|)"]])

  # Store one Cox estimate per repo. This becomes the outcome in the repo-level rq3 models.
  data.frame(
    repo = repo_name,
    repo_key = repo_key,
    status = "fit",
    reason = paste(unique(fit_warnings), collapse = " | "),
    n_rows = nrow(repo_rows),
    weighted_lines = sum(repo_rows$weight),
    ai_lines = ai_lines,
    human_lines = human_lines,
    ai_events = sum(ai_rows$event * ai_rows$weight),
    human_events = sum(human_rows$event * human_rows$weight),
    log_hazard = format_nullable_number(coef_row[["coef"]]),
    hazard_ratio = hazard_ratio,
    ci_lower = format_nullable_number(ci_row[["lower .95"]]),
    ci_upper = format_nullable_number(ci_row[["upper .95"]]),
    # updating to pull the robust SE
    #se = format_nullable_number(coef_row[["se(coef)"]]),
    se = format_nullable_number(se_value),
    p_value = p_value,
    concordance = format_nullable_number(fit_summary$concordance[[1]]),
    classification = cox_label(p_value, hazard_ratio),
    stringsAsFactors = FALSE
  )
}

line_survival_by_repo <- split(line_survival, line_survival$repo_key)
rq3_cox_results <- map_dfr(names(line_survival_by_repo), ~ run_cox_for_repo(line_survival_by_repo[[.x]], .x))

rq3_cox_results %>% count(status, classification)
```
should say something like: 
1     fit ai_higher_termination_hazard 77
2     fit  ai_lower_termination_hazard 89
3     fit          no_clear_difference 14
4 skipped                      skipped  3



## Stage 2 random-effects meta-regression

### Data Prep

#### Build repo-level modeling dataset
```{r}
# making clean dataset for analysis 
analysis_repos <- repo_summary %>%
  distinct(repo_key, repo) %>%        # repo = the slug (owner/name)
  arrange(repo) %>% select(repo, repo_key)
```


```{r}
# adding one-off variables to dataset

## adding num_stars and primary_langugage from full_sample_repos
analysis_repos <- analysis_repos %>% left_join(full_sample_repos %>% select(repo, metadata_num_stars, metadata_primary_language), by = "repo")
# check to see if there's any repos with null values 
analysis_repos[is.na(analysis_repos$metadata_num_stars), ]
analysis_repos[is.na(analysis_repos$metadata_primary_language), ]

## adding binary flag for whether repo has AGENTS.md file
# creating regex pattern for search
agent_cfg_pattern <- regex(
  "(^|/)(agents\\.md)$",
  ignore_case = TRUE
)
# creating flag and adding to repo
analysis_repos <- analysis_repos %>%
  left_join(
    commit_files %>%
      filter(str_detect(filename, agent_cfg_pattern)) %>%
      group_by(repo_key) %>%
      # prefer a root-level file, else earliest-seen; take its first date 
      arrange(str_detect(filename, "/"), commit_date, .by_group = TRUE) %>%
      summarise(
        hasAGENTSmd = 1L,
        .groups = "drop"
      ),
    by = "repo_key"
  ) %>%
  mutate(hasAGENTSmd = replace_na(hasAGENTSmd, 0L))
# casting hasAGENTSmd flag to binary factor 
#analysis_repos <- analysis_repos %>% mutate(hasAGENTSmd = factor(hasAGENTSmd,levels = c(0, 1)))
# check to see if there's any repos with null values 
analysis_repos[is.na(analysis_repos$hasAGENTSmd), ]

# adding repo aggregate no review rate
analysis_repos <- analysis_repos %>% left_join(repo_pr_no_review_aggregate_metrics, by = "repo")
``` 


```{r}
# adding remaining metrics to main analysis table 
## adding agent_type_labor_concentration_churn and agent_type_labor_concentration_commits metrics
# operationalzing agent_type_labor_concentration_churn and agent_type_labor_concentration_commits ()
# HHI (Herfindahl-Hirschman Index) = sum of squared shares; here, each AI agent's share of the repo's AI labor (by churn or by commits)
# HHI value ranges from 1/k (work split evenly across k agents) to 1 (a single agent does everything). 
# Used over Gini because we have few discrete categories (<=6 agents) rather than a continuous size distribution, and HHI reads directly as "how concentrated is AI work in one tool"
agent_concentration <- commits %>%
  filter(date >= observation_start_date,
         date <= observation_cutoff_date,
         category == "ai_coauthored",          
         !(repo %in% excluded_repos),
         !(sha %in% outlier_commit_shas)) %>%   
  separate_rows(labels, sep = ";") %>%
  mutate(label = str_trim(str_to_lower(labels))) %>%
  filter(label %in% ai_labels) %>%
  group_by(repo_key, label) %>%
  summarise(
    agent_churn   = sum(coalesce(as.numeric(changed_lines), 0)),  
    agent_commits = n(),
    .groups = "drop_last"
  ) %>%
  summarise(
    agent_type_labor_concentration_churn   = sum((agent_churn   / sum(agent_churn))^2),
    agent_type_labor_concentration_commits = sum((agent_commits / sum(agent_commits))^2),
    .groups = "drop"
  )

# adding to analysis repos dataset
analysis_repos <- analysis_repos %>%
  left_join(agent_concentration, by = "repo_key")

# checking to see if any repos have NA value 
analysis_repos[is.na(analysis_repos$agent_type_labor_concentration_churn), ]
analysis_repos[is.na(analysis_repos$agent_type_labor_concentration_commits), ]


## adding num_maintainers and num_contributors to dataset
# num maintainers = the smallest set of (non-bot) contributors who, ranked by commit count, together account for >=80% of the repo's commits (Pareto/core-developer rule)
# num contributors = the number of (non-bot) contributors who authored at least one commit to the repo in the observation window
# calculating number of maintainers per repo 
maintainer_counts <- commits %>%
  filter(date >= observation_start_date,
         date <= observation_cutoff_date,
         category != "other_bot", # excluding bot commits from total commit count 
         !(repo %in% excluded_repos),
         !(sha %in% outlier_commit_shas)) %>%
  mutate(
    contributor = coalesce(author_login, git_author_name)  # identity, login preferred
  ) %>%
  filter(!is.na(contributor)) %>%
  count(repo_key, contributor, name = "n_commits") %>%   # commits per contributor per repo
  arrange(repo_key, desc(n_commits)) %>%
  group_by(repo_key) %>%
  mutate(
    cum_share = cumsum(n_commits) / sum(n_commits),
    # rank at which cumulative share first reaches 0.80
    first_at_80 = which(cum_share >= 0.80)[1]
  ) %>%
  summarise(
    num_contributors = n(),
    num_maintainers  = first(first_at_80),   # smallest set covering >=80%
    .groups = "drop"
  )

# adding to analysis repos dataset
analysis_repos <- analysis_repos %>%
  left_join(maintainer_counts, by = "repo_key")

# checking to see if any repos have NA value 
analysis_repos[is.na(analysis_repos$num_maintainers), ]
analysis_repos[is.na(analysis_repos$num_contributors), ]


# calculating ai_source_code_share: fraction of AI changed lines landing in source_code files
ai_source_code_share <- commits %>%
  filter(date >= observation_start_date,
         date <= observation_cutoff_date,
         category == "ai_coauthored",
         !(repo %in% excluded_repos),
         !(sha %in% outlier_commit_shas)) %>%
  distinct(repo_key, sha) %>%
  inner_join(commit_file_classifications, by = c("repo_key", "sha" = "commit_sha")) %>%
  group_by(repo_key) %>%
  summarise(
    ai_source_code_share = sum(changed_lines[file_primary_role == "source_code"], na.rm = TRUE) /
                           sum(changed_lines, na.rm = TRUE),
    .groups = "drop"
  )


# calculating ai_contribution_concentration: gini of AI commit sizes (changed_lines per commit) 
# Gini coefficient (ai_contribution_concentration) values range from [0, 1]
# ai_contribution_concentration: Gini of per-commit AI churn within a repo, in [0,1]. 
# higher number = more concentrated (a few large AI commits account for most AI churn); lower number = more evenly spread (AI churn distributed across similar-sized commits).
ai_contribution_concentration <- commits %>%
  filter(date >= observation_start_date,
         date <= observation_cutoff_date,
         category == "ai_coauthored",
         !(repo %in% excluded_repos),
         !(sha %in% outlier_commit_shas)) %>%
  filter(!is.na(changed_lines)) %>%        # drop only true NAs (Gini errors on them)
  group_by(repo_key) %>%
  summarise(
    ai_contribution_concentration = ineq::Gini(changed_lines),
    .groups = "drop"
  )


# calculating test_file_change_share : n touches in test files / n touches all classified files 
test_file_change_share <- commit_file_classifications %>%
  semi_join(
    commits %>%
      filter(date >= observation_start_date,
             date <= observation_cutoff_date,
             category != "other_bot",          
             !(repo %in% excluded_repos),
             !(sha %in% outlier_commit_shas)),
    by = c("repo_key", "commit_sha" = "sha")
  ) %>%
  group_by(repo_key) %>%
  summarise(
    test_file_change_share = mean(file_primary_role == "test", na.rm = TRUE),
    .groups = "drop"
  )


# calculating SonarQube snapshot based metrics: note window-filtered
sonar_window <- sonarqube_snapshot_scans %>%
  filter(snapshot_date >= observation_start_date,
         snapshot_date <= observation_cutoff_date,
         !(repo %in% excluded_repos),
         scan_status == "ok", # drop failed scans
         !is.na(metric_ncloc)) %>%   # keep only usable snapshots)                     
  mutate(technical_debt_index = metric_sqale_index / metric_ncloc)

sonar_metrics <- sonar_window %>%
  group_by(repo_key) %>%
  arrange(snapshot_index, .by_group = TRUE) %>%
  summarise(
    technical_debt_index_start          = first(technical_debt_index),
    technical_debt_index_last           = last(technical_debt_index),
    technical_debt_index_avg            = mean(technical_debt_index, na.rm = TRUE),
    metric_duplicated_lines_density_start = first(metric_duplicated_lines_density),
    metric_duplicated_lines_density_last  = last(metric_duplicated_lines_density),
    metric_duplicated_lines_density_avg   = mean(metric_duplicated_lines_density, na.rm = TRUE),
    .groups = "drop"
  )


# adding to analysis repos dataset
analysis_repos <- analysis_repos %>%
  left_join(ai_source_code_share,          by = "repo_key") %>%
  left_join(ai_contribution_concentration, by = "repo_key") %>%
  left_join(test_file_change_share,        by = "repo_key") %>%
  left_join(sonar_metrics,                 by = "repo_key")


# Summarise commit activity to one row per repo. These variables describe how
# active the repo is, how much churn it has, and how much of that work is AI-coauthored.
# note excluding bot comits from commit counts as per standard practice
commit_activity_summary <- commits %>%
  filter(date >= observation_start_date,
         date <= observation_cutoff_date,
         category != "other_bot", # excluding bot commits from total commit count 
         !(repo %in% excluded_repos),
         !(sha %in% outlier_commit_shas)) %>%
  mutate(commit_month = substr(as.character(date), 1, 7),
         changed_lines = coalesce(as.numeric(changed_lines), 0)) %>%
  group_by(repo_key, repo) %>%
  summarise(
    total_commits_in_window = n(),
    total_line_churn_in_window = sum(changed_lines, na.rm = TRUE),
    ai_commits_in_window = sum(category == "ai_coauthored", na.rm = TRUE),
    ai_line_churn_in_window = sum(changed_lines[category == "ai_coauthored"], na.rm = TRUE),
    ai_commit_share = ratio_or_na(ai_commits_in_window, total_commits_in_window),
    ai_line_churn_share = ratio_or_na(ai_line_churn_in_window, total_line_churn_in_window),
    .groups = "drop"
  ) %>%
  mutate(
    log1p_total_commits_in_window = log1p_or_na(total_commits_in_window),
    log1p_total_line_churn_in_window = log1p_or_na(total_line_churn_in_window))


# adding to analysis repos dataset
analysis_repos <- analysis_repos %>%
  left_join(commit_activity_summary %>%  select(repo_key, total_commits_in_window,total_line_churn_in_window,ai_commits_in_window, ai_line_churn_in_window, ai_commit_share, ai_line_churn_share),by = "repo_key")


#deleting all the helper tables once metrics have been added to the main table
rm(ai_source_code_share, ai_contribution_concentration, test_file_change_share, sonar_window, sonar_metrics, commit_activity_summary, agent_concentration); gc()
```

building stage 2 dataset 
```{r}
# join per-repo Stage 1 effect (gap + SE + event counts) onto the repo factors
stage2_data <- rq3_cox_results %>%
  filter(status == "fit") %>%
  transmute(
    repo, repo_key,
    yi  = as.numeric(log_hazard),     # per-repo AI-vs-human log hazard ratio (the gap)
    sei = as.numeric(se),             # robust SE of that gap
    ai_events    = as.numeric(ai_events),
    human_events = as.numeric(human_events)
  ) %>%
  left_join(analysis_repos, by = c("repo", "repo_key"))
```



#### Pre-modeling data checks


```{r}
### Inspect the distributions of each variable, look for outliers

################################
##### ai_line_churn_share  #####
################################
summary(stage2_data$ai_line_churn_share)
ggplot(stage2_data, aes(x = ai_line_churn_share)) + 
  geom_histogram()
# moderate right-skew not too bad probably better to keep as-is


############################
##### ai_commit_share  #####
############################
summary(stage2_data$ai_commit_share)
ggplot(stage2_data, aes(x = ai_commit_share)) + 
  geom_histogram()
ggplot(stage2_data, aes(x = ai_commit_share, y = yi)) +
  geom_point(alpha = 0.5, color = "#F58F29") +
  geom_smooth(method = "lm", se = FALSE, color = "#59A14F") +
  labs(title = "gap vs ai_commit_share", y = "gap") + theme_minimal()


##########################
##### no_review_rate  #####
##########################
stage2_data$no_review_rate
summary(stage2_data$no_review_rate)
ggplot(stage2_data, aes(x = no_review_rate)) + 
  geom_histogram()
# most repos have a pretty low no merge rate (3rd quartile = 0.15) but there are some outliers
ggplot(stage2_data, aes(x = no_review_rate, y = yi)) +
  geom_point(alpha = 0.5, color = "#F58F29") +
  geom_smooth(method = "lm", se = FALSE, color = "#59A14F") +
  labs(title = "gap vs no_review_rate", y = "gap") + theme_minimal()



# TODO go through and finish inspecting the rest of the variables 
# <ONLY DO THIS FOR THE MODELS INCLUDED IN THE COVARIATE LIST NOT ALL OF THEM>
# mods = technical_dept_index_start + num_contributors + ai_source_code_share + test_file_change_share + no_review_rate


#########################################
##### ai_contribution_concentration  #####
#########################################
stage2_data$ai_contribution_concentration

summary(stage2_data$ai_line_churn_share)
ggplot(stage2_data, aes(x = ai_line_churn_share)) + 
  geom_histogram()
ggplot(stage2_data, aes(x = ai_line_churn_share)) + 
  geom_histogram() +
  scale_x_log10()

#########################################
##### test_file_change_share  #####
#########################################
stage2_data$test_file_change_share

summary(stage2_data$ai_line_churn_share)
ggplot(stage2_data, aes(x = ai_line_churn_share)) + 
  geom_histogram()
ggplot(stage2_data, aes(x = ai_line_churn_share)) + 
  geom_histogram() +
  scale_x_log10()


#########################################
##### technical_debt_index_start  #####
#########################################
stage2_data$technical_debt_index_start


summary(stage2_data$ai_line_churn_share)
ggplot(stage2_data, aes(x = ai_line_churn_share)) + 
  geom_histogram()
ggplot(stage2_data, aes(x = ai_line_churn_share)) + 
  geom_histogram() +
  scale_x_log10()

#########################################
##### ai_source_code_share  #####
#########################################
stage2_data$ai_source_code_share

summary(stage2_data$ai_line_churn_share)
ggplot(stage2_data, aes(x = ai_line_churn_share)) + 
  geom_histogram()
ggplot(stage2_data, aes(x = ai_line_churn_share)) + 
  geom_histogram() +
  scale_x_log10()

#########################################
##### hasAGENTSmd (binary) #####
#########################################
table(stage2_data$hasAGENTSmd)
ggplot(stage2_data, aes(x = factor(hasAGENTSmd), y = yi)) +
  geom_boxplot(fill = "#8789C0") +
  labs(title = "gap by hasAGENTSmd", x = "hasAGENTSmd (0/1)", y = "gap") + theme_minimal()

###########################################
##### metadata_primary_language  (categorical)      #####
###########################################
table(stage2_data$metadata_primary_language, useNA = "ifany")
# TODO: many levels; bin < rq3_min_language_repos into "other" via collapse_languages(). 
# note this factor is expensive in model (one slot per level) 

```

```{r}

stage2_model_data <- stage2_data %>%
  mutate(hasAGENTSmd = as.integer(as.character(hasAGENTSmd))) %>%   # clean 0/1
  # rma needs a valid gap + positive SE; drop rows missing any chosen covariate
  filter(!is.na(yi), !is.na(sei), sei > 0) %>%
  filter(ai_events >= 10, human_events >= 10) %>% # excluding repos that don't have at least 10 events in both groups 
  tidyr::drop_na(ai_line_churn_share, ai_contribution_concentration,
                 test_file_change_share, technical_debt_index_start,
                 ai_source_code_share, hasAGENTSmd)

# report eligible vs modeled so a silent drop from missing covariates is visible
n_eligible <- sum(!is.na(stage2_data$yi) & !is.na(stage2_data$sei) & stage2_data$sei > 0)
cat("eligible repos (valid gap + se):", n_eligible, "\n")
cat("repos in model:", nrow(stage2_model_data), "\n")
cat("dropped for missing covariate values:", n_eligible - nrow(stage2_model_data), "\n")


cat("\nStage 1 per-repo Cox hazard classification counts:\n")
rq3_cox_results %>%
  count(status, classification, name = "repos") %>%
  arrange(status, classification) %>%
  as.data.frame() %>%
  print(row.names = FALSE)

cat("\nStage 2 modeled repo hazard classification counts:\n")
rq3_cox_results %>%
  semi_join(stage2_model_data %>% select(repo_key), by = "repo_key") %>%
  count(status, classification, name = "repos") %>%
  arrange(status, classification) %>%
  as.data.frame() %>%
  print(row.names = FALSE)
```

#### Trying to determine which factors to include in rq3 model

```{r}
# Yi-vs-Vis Plots
library(tidyr)
candidate_mods <- c("metadata_num_stars", "hasAGENTSmd", "no_review_rate",
"agent_type_labor_concentration_churn", "agent_type_labor_concentration_commits",
"num_contributors", "num_maintainers", "ai_source_code_share",
"ai_contribution_concentration", "test_file_change_share",
"technical_debt_index_start", "technical_debt_index_last", "technical_debt_index_avg",
"metric_duplicated_lines_density_start", "metric_duplicated_lines_density_last",
"metric_duplicated_lines_density_avg", "total_commits_in_window",
"total_line_churn_in_window", "ai_commits_in_window", "ai_line_churn_in_window",
"ai_commit_share", "ai_line_churn_share")



# scatter of the gap against every candidate, with a loess + linear line
# look for: shapeless cloud (flat, no signal) vs a trend or curve (worth modeling)
stage2_model_data %>%
  select(yi, all_of(candidate_mods)) %>%
  pivot_longer(-yi, names_to = "moderator", values_to = "value") %>%
  ggplot(aes(value, yi)) +
  geom_point(alpha = 0.4, color = "#F58F29") +
  geom_smooth(method = "loess", se = FALSE, color = "#8789C0") +
  geom_smooth(method = "lm", se = FALSE, color = "#59A14F", linetype = 2) +
  facet_wrap(~ moderator, scales = "free_x") +
  labs(y = "per-repo gap (yi)") + theme_minimal()



# version with lagged skwewd vars
stage2_model_data %>%
  transmute(
    yi,
    log_total_commits = log1p(total_commits_in_window),
    log_total_churn   = log1p(total_line_churn_in_window),
    log_ai_commits    = log1p(ai_commits_in_window),
    log_ai_churn      = log1p(ai_line_churn_in_window),
    log_num_contributors = log1p(num_contributors),
    log_num_maintainers  = log1p(num_maintainers),
    log_ai_line_churn_in_window = log1p(ai_line_churn_in_window),
    log_technical_debt_index_start = log1p(technical_debt_index_start),
    log_metric_duplicated_lines_density_last = log1p(metric_duplicated_lines_density_last),
    log_stars         = log1p(metadata_num_stars)
  ) %>%
  pivot_longer(-yi, names_to = "moderator", values_to = "value") %>%
  ggplot(aes(value, yi)) +
  geom_point(alpha = 0.4, color = "#F58F29") +
  geom_smooth(method = "loess", se = FALSE, color = "#8789C0") +
  geom_smooth(method = "lm", se = FALSE, color = "#59A14F", linetype = 2) +
  facet_wrap(~ moderator, scales = "free_x") +
  labs(y = "per-repo gap (yi)", title = "gap vs LOGGED count moderators") +
  theme_minimal()
```


```{r}
## LASSO over the full candidate set to screen which moderators carry signal
set.seed(1)

# building the predictor matrix
# logging the skewed count/churn vars so LASSO sees them
lasso_df <- stage2_model_data %>%
  mutate(
    log_total_commits   = log1p(total_commits_in_window),
    log_total_churn     = log1p(total_line_churn_in_window),
    log_ai_commits      = log1p(ai_commits_in_window),
    log_ai_churn        = log1p(ai_line_churn_in_window),
    log_num_contributors= log1p(num_contributors),
    log_num_maintainers = log1p(num_maintainers),
    log_stars           = log1p(metadata_num_stars),
    primary_language_grouped = collapse_languages(metadata_primary_language,
                                                  min_repos = rq3_min_language_repos)
  )

# the full candidate set (logged counts + raw bounded vars + binary + language)
predictor_vars <- c(
  "log_stars", "hasAGENTSmd", "no_review_rate",
  "agent_type_labor_concentration_churn", "agent_type_labor_concentration_commits",
  "log_num_contributors", "log_num_maintainers",
  "ai_source_code_share", "ai_contribution_concentration", "test_file_change_share",
  "technical_debt_index_start", "technical_debt_index_last", "technical_debt_index_avg",
  "metric_duplicated_lines_density_start", "metric_duplicated_lines_density_last",
  "metric_duplicated_lines_density_avg",
  "log_total_commits", "log_total_churn", "log_ai_commits", "log_ai_churn",
  "ai_commit_share", "ai_line_churn_share",
  "primary_language_grouped"
)

# model.matrix expands the language factor into dummies and standardizes happens in glmnet
X <- model.matrix(~ . - 1, data = lasso_df[, predictor_vars])
y <- lasso_df$yi
# weight each repo by its precision so LASSO respects the meta-analytic structure
w <- 1 / (lasso_df$sei^2 + rq3_rma$tau2) ## TODO FIX THIS LINE where is rq3_rma

fit_path <- glmnet(X, y, weights = w, alpha = 1, standardize = TRUE)

# which variables enter FIRST as the penalty relaxes = strongest signal
# read left-to-right: the order variables become non-zero is the ranking
plot(fit_path, xvar = "lambda", label = TRUE)

# table with ranking values: for each variable, the largest lambda at which it's non-zero
# (bigger = enters earlier = stronger)
betas <- as.matrix(fit_path$beta)
entry_lambda <- apply(betas, 1, function(b) {
  nz <- which(b != 0)
  if (length(nz) == 0) NA else fit_path$lambda[min(nz)]
})
tibble(variable = rownames(betas), enters_at_lambda = entry_lambda) %>%
  filter(!is.na(enters_at_lambda)) %>%
  arrange(desc(enters_at_lambda)) %>%
  print(n = Inf)
```


```{r}
## Descriptive: do repos with more of these practices show higher overall survival?
## NOTE: this is the LEVEL (avg survival), not the AI-vs-human gap — exploratory context
# pulling per-repo overall survival rate from repo_summary
survival_explore <- analysis_repos %>%
  left_join(repo_summary %>% select(repo_key, survival_rate), by = "repo_key")

# binning the continuous practices into terciles (low/med/high) so patterns are visible
survival_binned <- survival_explore %>%
  transmute(
    survival_rate,
    `Stars`                              = ntile(metadata_num_stars, 3),
    `Agent labor concentration (churn)`  = ntile(agent_type_labor_concentration_churn, 3),
    `Agent labor concentration (commits)`= ntile(agent_type_labor_concentration_commits, 3),
    `No Review Rate`                     = ntile(no_review_rate, 3),
    `Num contributors`                   = ntile(num_contributors, 3),
    `Num maintainers`                    = ntile(num_maintainers, 3),
    `AI source-code share`               = ntile(ai_source_code_share, 3),
    `AI contribution concentration`      = ntile(ai_contribution_concentration, 3),
    `Test file change share`             = ntile(test_file_change_share, 3),
    `Technical debt (start)`             = ntile(technical_debt_index_start, 3),
    `Technical debt (last)`              = ntile(technical_debt_index_last, 3),
    `Technical debt (avg)`               = ntile(technical_debt_index_avg, 3),
    `Duplicated density (start)`         = ntile(metric_duplicated_lines_density_start, 3),
    `Duplicated density (last)`          = ntile(metric_duplicated_lines_density_last, 3),
    `Duplicated density (avg)`           = ntile(metric_duplicated_lines_density_avg, 3),
    `Total commits`                      = ntile(total_commits_in_window, 3),
    `Total line churn`                   = ntile(total_line_churn_in_window, 3),
    `AI commits`                         = ntile(ai_commits_in_window, 3),
    `AI line churn`                      = ntile(ai_line_churn_in_window, 3),
    `AI commit share`                    = ntile(ai_commit_share, 3),
    `AI churn share`                     = ntile(ai_line_churn_share, 3)
  )

# survival by each binned practice
survival_binned %>%
  pivot_longer(-survival_rate, names_to = "practice", values_to = "tercile") %>%
  filter(!is.na(tercile)) %>%
  ggplot(aes(factor(tercile), survival_rate)) +
  geom_boxplot(fill = "#8789C0") +
  facet_wrap(~ practice, scales = "free", ncol = 4) +     # 3 per row
  labs(x = "tercile (1 = low, 3 = high)", y = "repo avg survival rate",
       title = "Overall survival by practice tercile") +
  theme_minimal()

```


#### Building Model
```{r}
ggplot(stage2_model_data, aes(x=num_maintainers, y=no_review_rate)) + geom_point()
ggplot(stage2_model_data, aes(x = num_maintainers, y = yi)) +
  geom_point(alpha = 0.5, color = "#F58F29") +
  geom_smooth(method = "lm", se = FALSE, color = "#59A14F") +
  labs(title = "gap vs no_review_rate", y = "gap") + theme_minimal()
```


```{r}
# fitting model random-effects meta-regression 
# yi = per-repo gap, sei = its SE. rma weights each repo by 1/(sei^2 + tau^2)
# internally (do NOT pass weights). REML estimates tau^2; knha = small-sample CIs
# mods = technical_dept_index_start + num_contributors + ai_source_code_share + test_file_change_share +  ai_line_churn_share + no_review_rate
rq3_rma <- rma(
  yi   = yi,
  sei  = sei,
  mods = ~ technical_debt_index_start + 
    num_contributors +
    ai_source_code_share  + 
    test_file_change_share + 
    ai_line_churn_share + 
    no_review_rate,
  data   = stage2_model_data,
  method = "REML",
  test   = "knha")
summary(rq3_rma)

# tau^2 = residual between-repo variance in gap; I^2 = real heterogeneity share;
# R^2 = % of true between-repo variance the moderators explain (i.e., the explanatory power of the model)
c(tau2 = rq3_rma$tau2, I2 = rq3_rma$I2, R2 = rq3_rma$R2)

# coefficients on the hazard-ratio scale. HR_mult > 1: widens AI penalty; < 1: narrows
rq3_rma_table <- tibble(
  moderator = rownames(rq3_rma$beta),
  est_logHR = as.numeric(rq3_rma$beta),
  HR_mult   = exp(as.numeric(rq3_rma$beta)),
  ci_low    = exp(rq3_rma$ci.lb),
  ci_high   = exp(rq3_rma$ci.ub),
  p         = rq3_rma$pval
)
rq3_rma_table
```
Interpretation: 
Across 177 repositories, we find statistically significant evidence that the AI-vs-human maintenance gap depends on repository characteristics rather than varying at random (F(6,170) = 2.86, p = .011), with the rate of merging code without review emerging as the strongest individual predictor (HR = 1.75 per SD, 95% CI [1.01, 3.04], p = .046): repos that more often skip review have AI code that gets churned out faster relative to human code. Technical debt and limited test infrastructure show non-significant effects in the same direction (HR = 1.47 and 1.90, both p > .10), pointing to a coherent picture in which weaker engineering safeguards widen the AI penalty.
The fact that the model as a whole is significant while only one individual predictor reaches significance suggests two drivers: 
(1) Missing covariates: Our covariate set captures part of the signal but is likely incomplete — the characteristics driving the remaining heterogeneity may be finer-grained or unmeasured here
(2) Underpowered model: With 177 repositories individual moderator effects of the magnitude observed here (e.g., HR = 1.47 for technical debt) may be real but cannot be reliably distinguished from chance at this sample size. Confirming these directional effects would require a larger sample

```{r}
## Diagnostics
# 1. collinearity (rma won't warn): parallel lm -> VIF. >5 moderate, >10 high.
car::vif(lm(yi ~ technical_debt_index_start + num_contributors + ai_source_code_share  + test_file_change_share + ai_line_churn_share + no_review_rate,data = stage2_model_data))
# all VIF < 2 good

# 2. influential repos (huge-SE near-separation fits, size spread): any one repo driving it?
rq3_influence <- influence(rq3_rma)
plot(rq3_influence)
rq3_influence$inf[rq3_influence$inf$infl == "*", ]
# no influential repos

# 3. residual normality
qqnorm(residuals(rq3_rma, type = "rstandard")); qqline(residuals(rq3_rma, type = "rstandard"))
# clean qq plot
```

```{r}
# checking to see if the model is underpowered 
# how big an effect can we detect? simulate moderators with known effects,
set.seed(1)
detect_rate <- function(true_beta, n_sim = 200) {
  hits <- replicate(n_sim, {
    x <- rnorm(nrow(stage2_model_data))                 
    tau <- sqrt(rq3_rma$tau2)
    yi_sim <- true_beta * x + rnorm(nrow(stage2_model_data), 0, tau) +
              rnorm(nrow(stage2_model_data), 0, stage2_model_data$sei)
    fit <- rma(yi = yi_sim, sei = stage2_model_data$sei, mods = ~ x,
               method = "REML", test = "knha")
    fit$pval[2] < 0.05
  })
  mean(hits)
}
# detection rate (power) for per-SD effects of these sizes on the log-HR scale:
tibble(
  beta_logHR = c(0.1, 0.2, 0.3, 0.5),
  HR_per_SD  = exp(c(0.1, 0.2, 0.3, 0.5)),
  power      = sapply(c(0.1, 0.2, 0.3, 0.5), detect_rate)
)
```