
```{r setup, include=FALSE}
# loading libraries
# general use libraries
# library(tidyverse)
# library(RMySQL)
library(DBI)
library(readr)
library(ggplot2)
library(lubridate)
library(gt)
library(ggvenn)
library(ggridges)
# library(hrbrthemes)
library(gridExtra)
library(ggpubr)
library(tidyr)
library(purrr)
library(tibble)
library(numbers)
library(car)
library(GGally)
library(dotenv)
# libraries for survival curves
library(survival)
library(survminer)
library(broom)
library(xml2)
library(survey) # for cluster-robust chi-squared test with Rao-Scott correction
library(rcompanion) # for calculating Cramer's V of chi-squared tests
library(duckdb) # for handling big data
library(showtext) # for importing and using custom fonts
library(sysfonts)
library(systemfonts)
library(svglite) # for saving plots as svg to preserve custom font
library(lme4) # Poisson GLMMs
library(DHARMa) # GLMM residual diagnostics
library(ineq) # for calculating gini coefficient
library(tidyverse)
library(dplyr) # putting dplyr at the end because for some reason it misbehaves otherwise
library(glmnet) # for LASSO regression 
library(metafor)

# if dplyr is fighting with duckdb when trying to run Step 1 below, restart R session & fully quit R -> run this in console: install.packages(c("dbplyr", "duckdb", "dplyr")) -> then try rerunning this chunk
# or straight in terminal: pkill -f rsession -> Rscript -e 'install.packages(c("dbplyr", "duckdb", "dplyr"), repos="https://cloud.r-project.org")'
knitr::opts_chunk$set(echo = TRUE)
find_repo_root <- function(start = getwd()) {
  cur <- normalizePath(start, mustWork = TRUE)

  repeat {
    if (dir.exists(file.path(cur, "csv_full")) &&
        file.exists(file.path(cur, "filtered_full_sample_repos.csv"))) {
      return(cur)
    }

    parent <- dirname(cur)
    if (identical(parent, cur)) {
      stop("Could not find repo root containing csv_full/")
    }

    cur <- parent
  }
}

repoRoot <- find_repo_root()
csv_full_dir <- file.path(repoRoot, "csv_full", "")
```

# rq2 

# Loading data
Loading data workflow:

```{r}
## Analysis constants (RUN EVERY SESSION)

#filters established and verified during prelim exploration (see Full_Dataset_Exploration.Rmd doc)
outlier_commit_shas <- c("2b717338b0e02663dfce5142bd636db0b1d3958c","a3f5c98b59b7892866a49ac22ed91cd8c37865d8","038ca6696d9b821b216b8ab06c8dd48aac7750ae","cea8cd84957f4801f3e9ea1f0997fedfbe3c5e35","dc01a470926d8ee353d7627171e0d491bb8e6386","00314a6c4a2a1c422a4cf9a57b15efca1ad0e146","10763bbae6614c3dc569da892ce6453afd0143f4","ec3a5ff1142670bed5f43969581ecd179d2b8d50","66b199403e65ee29364a25cd7b7596e679030e7f","57960940f9a0b6cd9caf0f92c3592be0cad588d5","23a42d5d956b04c49276240875d8dd329eb44642","0fb2e4b1968f9db4f1cf36fe89505f536e82bc94","3760f7ddeba3cd857a4068a5c08df2800c6950ec")

# these are repos we're excluding: statsig-io/docs - because it was deleted so we couldn't collect full data, the rest because according to the data they don't have any ai_coauthored commits
# github/copilot-cli in particular was a separate discover later on, its a cli tool with no source code so we exclueded it for that reason 
excluded_repos <- c("statsig-io/docs", "Daylily-Informatics/daylily-ephemeral-cluster", "TypedDevs/bashunit", "tscircuit/contribution-tracker", "tscircuit/tscircuit.com", "github/copilot-cli")

observation_start_date <- as.POSIXct("2025-05-01 00:00:00", format = "%Y-%m-%d %H:%M:%S", tz = "UTC")
observation_cutoff_date <- as.POSIXct("2026-05-31 23:59:59", format = "%Y-%m-%d %H:%M:%S", tz = "UTC") 
observation_start_year_month <- "2025-05"
observation_cutoff_year_month <- "2026-05" 
```

##  Model 1A: Binomial mixed-effects regression of corrective commit rate (Lagged AI code share)

Question: Does the overall corrective commit rate rise as the previous month's Agentic code-share snapshot increases?
```{r}
rq2_panel_csv_path <- paste0(csv_full_dir, "rq2_repo_week_panel.csv")

rq2_min_total_commits <- 3L

rq2_last_finite <- function(values) {
  values <- values[!is.na(values) & is.finite(values)]
  if (length(values) == 0) {
    return(NA_real_)
  }
  tail(values, 1)
}

rq2_week_panel_raw <- read_csv(rq2_panel_csv_path, show_col_types = FALSE)
rq2_week_column <- intersect(c("week", "week_date", "repo_week"), names(rq2_week_panel_raw))
if (length(rq2_week_column) == 0) {
  stop("rq2_repo_week_panel.csv must contain one of these week columns: week, week_date, repo_week")
}
rq2_week_column <- rq2_week_column[[1]]

rq2_week_panel <- rq2_week_panel_raw %>%
  mutate(
    week_date = as.Date(as.character(.data[[rq2_week_column]])),
    repo_month = format(week_date, "%Y-%m"),
    month_start = as.Date(paste0(repo_month, "-01")),
    ai_share = as.numeric(ai_share),
    total_commits = as.numeric(total_commits),
    corrective_commits = as.numeric(corrective_commits),
    ai_total_commits = as.numeric(ai_total_commits),
    human_total_commits = as.numeric(human_total_commits),
    human_corrective_commits = as.numeric(human_corrective_commits)
  ) %>%
  filter(repo_month >= observation_start_year_month,
         repo_month <= observation_cutoff_year_month,
         !(repo %in% excluded_repos))

rq2_panel_lagged_code_share <- rq2_week_panel %>%
  arrange(repo_key, week_date) %>%
  group_by(repo_key, repo, repo_month, month_start) %>%
  summarise(
    # Use the last observed weekly code-share snapshot in the month, then lag it
    # by one month below. 
    ai_share_month = rq2_last_finite(ai_share),
    total_commits = sum(total_commits, na.rm = TRUE),
    corrective_commits = sum(corrective_commits, na.rm = TRUE),
    ai_commits = sum(ai_total_commits, na.rm = TRUE),
    human_total_commits = sum(human_total_commits, na.rm = TRUE),
    human_corrective_commits = sum(human_corrective_commits, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(repo_key, month_start) %>%
  group_by(repo_key) %>%
  mutate(
    living_line_ai_share_lag1 = lag(ai_share_month, 1),
    ai_commit_share = if_else(total_commits > 0, ai_commits / total_commits, NA_real_),
    obs_id_base = row_number()
  ) %>%
  ungroup()

rq2_panel_lagged_code_share %>%
  summarise(
    repo_months = n(),
    repos = n_distinct(repo_key),
    months = n_distinct(repo_month),
    repo_months_with_lagged_share = sum(!is.na(living_line_ai_share_lag1)),
    mean_lagged_agentic_code_share = mean(living_line_ai_share_lag1, na.rm = TRUE)
  )

rq2_prepare_glmer_data <- function(panel, successes_col, trials_col, size_col = trials_col) {
  panel %>%
    filter(!is.na(living_line_ai_share_lag1),
           is.finite(living_line_ai_share_lag1),
           !is.na(.data[[successes_col]]),
           !is.na(.data[[trials_col]]),
           !is.na(.data[[size_col]]),
           is.finite(.data[[successes_col]]),
           is.finite(.data[[trials_col]]),
           is.finite(.data[[size_col]]),
           total_commits >= rq2_min_total_commits,
           .data[[trials_col]] > 0,
           .data[[size_col]] > 0,
           .data[[successes_col]] >= 0,
           .data[[successes_col]] <= .data[[trials_col]],
           !is.na(repo_key),
           nzchar(repo_key)) %>%
    mutate(
      successes = round(.data[[successes_col]]),
      trials = round(.data[[trials_col]]),
      failures = trials - successes,
      log_trials = log(as.numeric(.data[[size_col]])),
      repo_key = factor(repo_key)
    ) %>%
    filter(successes >= 0,
           failures >= 0,
           trials > 0,
           is.finite(log_trials)) %>%
    arrange(repo_key, month_start) %>%
    mutate(obs_id = factor(row_number()))
}

rq2_corrective_model_data <- rq2_prepare_glmer_data(
  rq2_panel_lagged_code_share,
  successes_col = "corrective_commits",
  trials_col = "total_commits",
  size_col = "total_commits"
)

rq2_corrective_lagged_code_share_fit <- glmer(
  cbind(successes, failures) ~
    living_line_ai_share_lag1 +
    log_trials +
    (1 | repo_key) +
    (1 | obs_id),
  data = rq2_corrective_model_data,
  family = binomial(link = "logit"),
  control = glmerControl(optimizer = "bobyqa")
)

summary(rq2_corrective_lagged_code_share_fit)
isSingular(rq2_corrective_lagged_code_share_fit)
VarCorr(rq2_corrective_lagged_code_share_fit)

# diagnostics 
# --- DIAGNOSTIC: Convergence / singular fit ---
# Want to see: FALSE, no convergence warnings.
# Don't want: TRUE (repo variance ~0) or "failed to converge".
isSingular(rq2_corrective_lagged_code_share_fit)

# --- DIAGNOSTIC: Distributional fit ---
# Test/plot: DHARMa::simulateResiduals(re.form = NULL) -> plot.
# Want to see: QQ on the diagonal, non-significant KS, flat quantile lines.
# Don't want: KS deviation / curved quantile lines.
rq2_corrective_sim <- DHARMa::simulateResiduals(
  rq2_corrective_lagged_code_share_fit,
  simulateREs = "user-specified",
  re.form = NULL
)
plot(rq2_corrective_sim)

# --- DIAGNOSTIC: Overdispersion ---
# Test: Pearson chi-square / residual df (same as the GLM version, with GLMM-correct df)
# Want to see: ratio ~1.
# Don't want: >1 — SEs too small; refit with glmer.nb (same formula).
rq2_corrective_rp <- residuals(rq2_corrective_lagged_code_share_fit, type = "pearson")
rq2_corrective_rdf <- nobs(rq2_corrective_lagged_code_share_fit) -
  length(fixef(rq2_corrective_lagged_code_share_fit)) -
  length(getME(rq2_corrective_lagged_code_share_fit, "theta"))
c(chisq = sum(rq2_corrective_rp^2),
  ratio = sum(rq2_corrective_rp^2) / rq2_corrective_rdf,
  rdf = rq2_corrective_rdf,
  p = pchisq(sum(rq2_corrective_rp^2), df = rq2_corrective_rdf, lower.tail = FALSE))

# --- DIAGNOSTIC: Zero-success check ---
# Grouped-binomial model: observed repo-months with zero corrective commits vs
# expected zero-success rows under the fitted binomial probabilities.
rq2_corrective_successes <- rq2_corrective_model_data$successes
rq2_corrective_trials <- rq2_corrective_model_data$trials
rq2_corrective_p_hat <- fitted(rq2_corrective_lagged_code_share_fit)
c(observed = sum(rq2_corrective_successes == 0),
  expected = sum(dbinom(0, size = rq2_corrective_trials, prob = rq2_corrective_p_hat)),
  ratio = sum(rq2_corrective_successes == 0) /
    sum(dbinom(0, size = rq2_corrective_trials, prob = rq2_corrective_p_hat)))
```


##  Model 1B: Binomial mixed-effects regression of human corrective commit rate (Lagged AI code share)

Question: Does the human-authored corrective commit rate rise as the previous month's Agentic code-share snapshot increases?
```{r}
rq2_human_corrective_model_data <- rq2_prepare_glmer_data(
  rq2_panel_lagged_code_share,
  successes_col = "human_corrective_commits",
  trials_col = "human_total_commits",
  size_col = "human_total_commits"
)

rq2_human_corrective_lagged_code_share_fit <- glmer(
  cbind(successes, failures) ~
    living_line_ai_share_lag1 +
    log_trials +
    (1 | repo_key) +
    (1 | obs_id),
  data = rq2_human_corrective_model_data,
  family = binomial(link = "logit"),
  control = glmerControl(optimizer = "bobyqa")
)

summary(rq2_human_corrective_lagged_code_share_fit)
isSingular(rq2_human_corrective_lagged_code_share_fit)
VarCorr(rq2_human_corrective_lagged_code_share_fit)

# diagnostics 
# --- DIAGNOSTIC: Convergence / singular fit ---
# Want to see: FALSE, no convergence warnings.
# Don't want: TRUE (repo variance ~0) or "failed to converge".
isSingular(rq2_human_corrective_lagged_code_share_fit)

# --- DIAGNOSTIC: Distributional fit ---
# Test/plot: DHARMa::simulateResiduals(re.form = NULL) -> plot.
# Want to see: QQ on the diagonal, non-significant KS, flat quantile lines.
# Don't want: KS deviation / curved quantile lines.
rq2_human_corrective_sim <- DHARMa::simulateResiduals(
  rq2_human_corrective_lagged_code_share_fit,
  simulateREs = "user-specified",
  re.form = NULL
)
plot(rq2_human_corrective_sim)

# --- DIAGNOSTIC: Overdispersion ---
# Test: Pearson chi-square / residual df (same as the GLM version, with GLMM-correct df)
# Want to see: ratio ~1.
# Don't want: >1 — SEs too small; refit with glmer.nb (same formula).
rq2_human_corrective_rp <- residuals(rq2_human_corrective_lagged_code_share_fit, type = "pearson")
rq2_human_corrective_rdf <- nobs(rq2_human_corrective_lagged_code_share_fit) -
  length(fixef(rq2_human_corrective_lagged_code_share_fit)) -
  length(getME(rq2_human_corrective_lagged_code_share_fit, "theta"))
c(chisq = sum(rq2_human_corrective_rp^2),
  ratio = sum(rq2_human_corrective_rp^2) / rq2_human_corrective_rdf,
  rdf = rq2_human_corrective_rdf,
  p = pchisq(sum(rq2_human_corrective_rp^2), df = rq2_human_corrective_rdf, lower.tail = FALSE))

# --- DIAGNOSTIC: Zero-success check ---
# Grouped-binomial model: observed repo-months with zero human corrective commits
# vs expected zero-success rows under the fitted binomial probabilities.
rq2_human_corrective_successes <- rq2_human_corrective_model_data$successes
rq2_human_corrective_trials <- rq2_human_corrective_model_data$trials
rq2_human_corrective_p_hat <- fitted(rq2_human_corrective_lagged_code_share_fit)
c(observed = sum(rq2_human_corrective_successes == 0),
  expected = sum(dbinom(0, size = rq2_human_corrective_trials, prob = rq2_human_corrective_p_hat)),
  ratio = sum(rq2_human_corrective_successes == 0) /
    sum(dbinom(0, size = rq2_human_corrective_trials, prob = rq2_human_corrective_p_hat)))
```


